---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# dotdot

This package contains 10 lines of code, it proposes an improved assignment using
the shorthand  `..`.

Think about the `..` as the `:` of the `:=` sign laid horizontally.

```{r}
library(dotdot)
x <- y <- iris
x$Sepal.Length[5] <- x$Sepal.Length[5] + 3
y$Sepal.Length[5] := .. + 3
identical(x,y)

z <- factor(letters[1:3])
levels(z) := c(.., "level4")
z
```

In case you've attached a package containing `:=` such as *`data.table`* or *`rlang`* you can use `dotdot_first()` to
make sure that our `:=` is not masked, the operator will still work fine in *`data.table`* and *`rlang`* as they only use it
for parsing but don't actually call it.

```{r}
library(data.table)
try(levels(z) := c(.., "level5")) # fails!
z
dotdot_first()
levels(z) := c(.., "level5") # works!
z
data <- as.data.table(head(iris))
data[,new_col := 3] # `:= ` works as if dotdot wasn't attached
data 
```
