---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# dotdot

Installation :

```{r}
devtools::install_github("moodymudskipper/dotdot")
```

This package proposes an improved assignment using the shorthand  `..`.


```{r}
library(dotdot)
x <- y <- iris
x$Sepal.Length[5] <- x$Sepal.Length[5] + 3
y$Sepal.Length[5] := .. + 3
identical(x,y)

z <- factor(letters[1:3])
levels(z) := c(.., "level4")
z
```

You can think about the `..` as the `:` of the `:=` symbol laid horizontally.

## Integration with *`data.table`*, *`tidyverse`* and other packages using `:=`

The operator `:=` is used by prominent packages *`data.table`* and *`rlang`*
 (mostly through *`tidyverse`* functions), but they only use it to parse expressions, 
 due to its convenient operator precedence. It's not actually called. 
 
Thus *`dotdot`* is completely *`tidyverse`* and *`data.table`* compatible, and
some adhoc adjustments were made so it even works when the latter are attached
after *`dotdot`*.

```{r}
library(data.table)
levels(z) := c(.., "level5")
z
data <- as.data.table(head(iris,2))
data[,new_col := 3] # `:= ` works as if dotdot wasn't attached
data 
```

An example of fine integration of the operator being used by *`dotdot`* and 
*`rlang`* through *`dplyr`*

```{r}
library(dplyr)
my_data_frame <- iris[3:5]
var = quo(Log.Petal.Width)
my_data_frame := .. %>% mutate(!!var := log(Petal.Width)) %>% head(2)
my_data_frame
```

In case you've attached another package containing `:=`, you can use `dotdot_first()` to
make sure that our `:=` is not masked (It seems to be rare enough though as I
couldn't find an example)

## Comparison with magrittr's `%<>%`

The package *`magrittr`* contains the operator `%<>%` which serves a similar role
to `:=`. Let's see how it is similar first, and then how it differs :

These calls have the same effect:

```{r,eval = FALSE}
iris$Sepal.Length %<>% log
iris$Sepal.Length %<>% log(.)
iris$Sepal.Length := log(..)
```

Those as well, but here we see *`magrittr`* is less compact and readable.

```{r,eval = FALSE}
iris$Sepal.Length[5] %<>% multiply_by(2) %>% add(3)
iris$Sepal.Length[5] %<>% {2*. + 3}
iris$Sepal.Length[5] := 2*.. + 3
```

Now for the differences, aside from compacity and readability :

* attaching *`magrittr`* means often masking functions likes `extract` or 
`set_names`. *`dotdot`* only exports its operator and the `dotdot_first` function.
* *`magrittr`* operators deal with environment in a way that is much less straightforward, so this won't work :

```{r}
library(magrittr)
test <- function(some_parameter) {
  some_parameter %<>% {as.character(substitute(.))}
  some_parameter
  }
x <- try(test(foo))
inherits(x,"try-error")
```

While this will work fine:

```{r}
test <- function(some_parameter) {
  some_parameter := as.character(substitute(..))
  some_parameter
  }
test(foo)
```

*  `:=` is also faster than `%<>%` , though these operations are fast anyway and not likely to be a bottleneck very often if ever:

```{r}
 b <- x <- y <- z <- 1
 microbenchmark::microbenchmark(
   base      = {b <- b + 1},
   dotdot    = {x := .. + 1},
   magrittr  = {y %<>% add(1)},
   magrittr2 = {z %<>% {. + 1}},
   times = 1e4
)
```

## Edge cases and good practice

`:=` is **NOT** meant to be a complete replacement of the `<-` operator, the latter is explicit in the absence of `..` , so more readable, is faster (though we're speaking microseconds), and won't clutter your `traceback()` when debugging. 
 
`:=` can be used several times in a statement like `z <- (x := .. + 1) + (y:= .. +1)` but it never makes sense to use it `:=` several times in an assignment such as `x := (y := .. + 2)` as all the `..` will be replaced by the name of the variable on the lhs of the first evaluated `:=` in any case.  It can even produce counter intuitive output, see below.


This is all good and explicit :

```
x <- 4
y <- 7
z <- (x := .. + 1) + (y:= .. +1)
x
y
z
```

But using several nested `:=` is unuseful and potentially confusing, here the
dots will be replaced by `x`, though one might have expected them to be replaced
by `y`.

```
x <- 4
y <- 7
x := (y := .. + 2) # same as `x <- (y := x + 2)` 
x
y
```

Good practice makes things unambiguous :

```
x <- 4
y <- 7
x <- (y := .. + 2)
x
y

x <- 4
y <- 7
x := (y <- .. + 2)
x
y
```

