---
title: "dotdot"
author: "Antoine Fabri"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This package proposes an enhanced assignment using the shorthand  `..`.

It's meant to :

* Be explicit about growing an object or overwriting it using its previous value
* Avoid repeating a variable name
* Save keystrokes/time/visual space/cognitive load

```{r}
library(dotdot)
x <- y <- iris
x$Sepal.Length[5] <- x$Sepal.Length[5] + 3
y$Sepal.Length[5] := .. + 3
identical(x,y)

z <- factor(letters[1:3])
levels(z) := c(.., "level4")
z
```

It can help to think about the `..` as the `:` of the `:=` symbol laid horizontally.

## Integration with *`data.table`*, *`tidyverse`* and other packages using `:=`

The operator `:=` is used by prominent packages *`data.table`* and *`rlang`*
 (mostly through *`tidyverse`* functions), but they only use it to parse expressions, 
 due to its convenient operator precedence. It's not actually called. 
 
Thus *`dotdot`* is completely *`tidyverse`* and *`data.table`* compatible, and
some adhoc adjustments were made so it works when the latter are attached
after *`dotdot`*.

In the rare case in which the user would attach another package containing `:=`,
`dotdot_first()` will move back *`dotdot`* at the first position of the search
path.

## Comparison with magrittr's `%<>%`

The package *`magrittr`* contains the operator `%<>%` which share important
similarities with `:=`. 

The calls `iris$Sepal.Length %<>% log()` and `iris$Sepal.Length := log(..)` 
indeed have the same effect.

The operator precedence is different however, which makes *`dotdot`* arguably
better suited outside of pipe chains, for instance
`iris$Sepal.Length[5] := 2*.. + 3` would have to be written
`iris$Sepal.Length[5] %<>% multiply_by(2) %>% add(3)` or 
`iris$Sepal.Length[5] %<>% {2*. + 3}` using *`magrittr`*.

Furthermore *`magrittr`* is more complex in the way it deals with environments
while *`dotdot`* makes straightforward substitutions, so `x := substitute(..)`
works and doesn't have a *`magrittr`* equivalent.
